---
title: "Seasonal hydrographs"
author: "Sam Bashevkin"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = 'docs',
      knit_root_dir = "../",
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(DroughtData)
library(ggplot2)
```

```{r}
d<-DroughtData::raw_hydro_1975_2021%>%
  mutate(Year=year(Date),
         WY=if_else(month(Date)>=10, Year+1, Year),
         doy=yday(Date),
         wdoy=case_when(
           leap_year(Year) & doy >=275 ~ doy -274,
           leap_year(Year) & doy <275 ~ doy + 92,
           !leap_year(Year) & doy >=274 ~ doy -273,
           !leap_year(Year) & doy <274 ~ doy + 92
         ))%>%
  group_by(WY)%>%
  arrange(wdoy)%>%
  mutate(across(c(InflowSacR, InflowYolo, InflowEast, InflowTotal, Outflow, Export), list(cum=~cumsum(.x), sum=~sum(.x))))%>%
  left_join(DroughtData::lt_regional%>%
              distinct(YearAdj, SVIndex, YearType, Drought)%>%
              arrange(Drought, YearAdj)%>%
              group_by(Drought)%>%
              mutate(Lag=YearAdj-lag(YearAdj, order_by = YearAdj))%>%
              ungroup()%>%
              mutate(Start=if_else(is.na(Lag) | Lag>1, TRUE, FALSE),
                     Series_ID=1:n(),
                     Series_ID=if_else(Start, Series_ID, NA_integer_),
                     Series_ID=as.integer(as.factor(Series_ID)))%>%
              fill(Series_ID, .direction="down")%>%
              group_by(Series_ID)%>%
              arrange(YearAdj)%>%
              mutate(Series_year=1:n(),
                     Series_year2=if_else(Series_year>=3, "3+", as.character(Series_year)),
                     Series_year2=factor(Series_year2, levels=c("1", "2", "3+")))%>%
              select(-Lag, -Start),
            by=c("WY"="YearAdj"))%>%
  filter(!is.na(Drought))

# Calculate wdoy for the first day of each month
wdoy_months<-d%>%
  mutate(Month=month(Date, label=T))%>%
  group_by(Month)%>%
  summarise(day_min=min(wdoy))%>%
  mutate(month_label=day_min+15)
```

Total inflow seasonal hydrograph by year
```{r}
ggplot(d, aes(x=doy, y=InflowTotal, color=Series_year2, group=WY))+
  geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  facet_grid(Drought~., scales="free_y")+
  scale_color_viridis_d()+
  coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()
```

Total inflow seasonal hydrograph by series_year
```{r}
ggplot(d, aes(x=wdoy, y=InflowTotal, color=Series_year2, group=Series_year2))+
  geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  facet_grid(Drought~., scales="free_y")+
  xlab("Water day of year")+
  scale_color_viridis_d()+
  scale_x_continuous(breaks=c(wdoy_months$day_min, wdoy_months$month_label), labels=c(rep("", 12), as.character(wdoy_months$Month)))+
  coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()+
  theme(axis.ticks.x = element_line(color = c(rep("black", 12), rep(NA, 12))))
```


Total inflow seasonal hydrograph by series_year, scaled by summed inflow over the year
```{r}
ggplot(d, aes(x=wdoy, y=InflowTotal/InflowTotal_sum, color=Series_year2, group=Series_year2))+
  geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  facet_grid(Drought~.)+
  xlab("Water day of year")+
  scale_color_viridis_d()+
  scale_x_continuous(breaks=c(wdoy_months$day_min, wdoy_months$month_label), labels=c(rep("", 12), as.character(wdoy_months$Month)))+
  coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()+
  theme(axis.ticks.x = element_line(color = c(rep("black", 12), rep(NA, 12))))
```

Total cumulative inflow seasonal hydrograph by series_year
```{r}
ggplot(d, aes(x=wdoy, y=InflowTotal_cum/InflowTotal_sum, color=Series_year2, group=WY))+
  geom_line()+
  facet_grid(Drought~., scales="free_y")+
  ylab("Cumulative proportional inflow")+
  xlab("Water day of year")+
  scale_color_viridis_d()+
  scale_x_continuous(breaks=c(wdoy_months$day_min, wdoy_months$month_label), labels=c(rep("", 12), as.character(wdoy_months$Month)))+
  coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()+
  theme(axis.ticks.x = element_line(color = c(rep("black", 12), rep(NA, 12))))
```

This shows most clearly how in dry years, after 2+ years of a drought, water managers are more conservative early in the water year, saving up whatever rainfall they can. But in wet years after 2+ years of a wet period, the reservoirs are full and much more water is released earlier in the water year. Neutral years show no pattern, as expected since they represent flip-flopping among year types.


Same plot as above, but smoothed for each series_year and drought category combo
```{r}
ggplot(d, aes(x=wdoy, y=InflowTotal_cum/InflowTotal_sum, color=Series_year2, group=Series_year2))+
  geom_smooth(method="gam", formula=y~s(x, bs="cs", k=25))+
  facet_grid(Drought~., scales="free_y")+
  ylab("Cumulative proportional inflow")+
  xlab("Water day of year")+
  scale_color_viridis_d()+
  scale_x_continuous(breaks=c(wdoy_months$day_min, wdoy_months$month_label), labels=c(rep("", 12), as.character(wdoy_months$Month)))+
  coord_cartesian(ylim=c(0, NA), expand = FALSE)+
  theme_bw()+
  theme(axis.ticks.x = element_line(color = c(rep("black", 12), rep(NA, 12))))
```

