---
title: "Zoop-salinity CSAMP SDM model"
author: "Sam Bashevkin"
date: "8/19/2022"
output: html_document
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = 'docs',
      envir = globalenv()
    )
    })
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

Load packages
```{r, message=FALSE}
require(dplyr)
require(zooper)
require(lubridate)
require(readr)
require(tidyr)
require(ggplot2)
require(sf)
require(readxl)
require(stringr)
require(mgcv)
require(purrr)
```

Load zoop data
```{r}
zoop_data<-Zoopsynther(Data_type="Community", Sources=c("EMP", "STN", "20mm", "FMWT"), Time_consistency = TRUE)
```


Read in zoop mass conversions
```{r}
zoop_mass_conversions<-read_excel("~/ZoopSynth/Data paper/Biomass conversions.xlsx", sheet="Micro and Meso-zooplankton")%>%
  mutate(Taxname=case_when(Taxname=="Sinocalanus"~"Sinocalanus doerrii", # Change to help this match to zoop data
                           TRUE ~ Taxname),
         Taxlifestage=paste(Taxname, Lifestage))%>%
  select(Taxlifestage, CarbonWeight_ug)
```

Read in zoop groupings
```{r}
zoop_groups<-read_csv("Data/zoopcrosswalk2.csv", col_types=cols_only(Taxlifestage="c", IBMR="c"))%>%
  distinct()
```


Load Mysid biomass data
```{r}
zoop_mysid<-read_excel("Data/1972-2020MysidBPUEMatrix.xlsx", # EMP
                       sheet="Mysid_BPUE_matrix_1972-2020", na = "NA",
                       col_types = c(rep("numeric", 4), "date", "text", "text", rep("text", 7), rep("numeric", 8)))%>%
  select(Date=SampleDate, Station=StationNZ, BPUE=`Hyperacanthomysis longirostris`)%>% # Only select Hyperacanthomysis longirostris
  mutate(Source="EMP")%>%
  bind_rows(read_csv("Data/FMWT STN 2007to2019 Mysid BPUE.csv", # FMWT/STN
                     col_types=cols_only(Station="c", SampleDate="c", Project="c", `Hyperacanthomysis longirostris`="d"))%>% 
              rename(Date=SampleDate, Source=Project, BPUE=`Hyperacanthomysis longirostris`)%>% # Only select Hyperacanthomysis longirostris
              mutate(Date=mdy(Date),
                     Station=recode(Station, MONT="Mont", HONK="Honk")))%>% #Get station names to match to main dataset
  mutate(BPUE_mysid=BPUE*1000, # Convert to ug
         Taxlifestage="Hyperacanthomysis longirostris Adult",
         SampleID=paste(Source, Station, Date),
         SizeClass="Macro")%>%
  select(SampleID, Taxlifestage, SizeClass, BPUE_mysid)
```


Start processing the zoop data
```{r}
zoop_data_mass<-zoop_data%>%
  mutate(Taxlifestage=str_remove(Taxlifestage, fixed("_UnID")))%>%
  filter(
    !(SizeClass=="Meso" & #eliminating species which are counted in meso and micro and retained better in the micro net from the meso calcs
        
        Taxlifestage%in%c("Asplanchna Adult", "Copepoda Larva","Cyclopoida Juvenile", "Eurytemora Larva", "Harpacticoida Undifferentiated",
                          "Keratella Adult", "Limnoithona Adult", "Limnoithona Juvenile", "Limnoithona sinenesis Adult", "Limnoithona tetraspina
                                    Adult", "Oithona Adult", "Oithona Juvenile", "Oithona davisae Adult", "Polyarthra Adult","Pseudodiaptomus Larva", 
                          "Rotifera Adult", "Sinocalanus doerrii Larva", "Synchaeta Adult", "Synchaeta bicornis Adult", "Trichocerca Adult")) &
      
      !(SizeClass=="Micro" &Taxlifestage%in%c("Cirripedia Larva", "Cyclopoida Adult", "Oithona similis")) & #removing categories better retained in meso net from micro net matrix
      Order!="Amphipoda" & # Remove amphipods
      (Order!="Mysida" | Taxlifestage=="Hyperacanthomysis longirostris Adult"))%>% #Only retain Hyperacanthomysis longirostris
  mutate(Taxlifestage=recode(Taxlifestage, `Synchaeta bicornis Adult`="Synchaeta Adult", # Change some names to match to biomass conversion dataset
                             `Pseudodiaptomus Adult`="Pseudodiaptomus forbesi Adult",
                             `Acanthocyclops vernalis Adult`="Acanthocyclops Adult"))%>%
  left_join(zoop_mass_conversions, by="Taxlifestage")%>% # Add biomass conversions
  left_join(zoop_mysid, by=c("SampleID", "Taxlifestage", "SizeClass"))%>% # Add mysid biomass
  left_join(zoop_groups, by="Taxlifestage")%>% # Add IBMR categories
  mutate(BPUE=if_else(Taxlifestage=="Hyperacanthomysis longirostris Adult", BPUE_mysid, CPUE*CarbonWeight_ug))%>% # Create 1 BPUE variable
  filter(!is.na(BPUE) & !is.na(Latitude) & !is.na(Longitude) & !is.na(SalSurf))%>% # Removes any data without BPUE, which is currently restricted to Decapod Larvae, and H. longirostris from STN. Also removes 20mm and EMP EZ stations without coordinates
  group_by(IBMR)%>%
  mutate(flag=if_else(all(c("Micro", "Meso")%in%SizeClass), "Remove", "Keep"))%>% # This and the next 2 lines are meant to ensure that all categories are consistent across the surveys. Since only EMP samples microzoops, only EMP data can be used for categories that include both micro and mesozoops.
  ungroup()%>%
  filter(!(flag=="Remove" & Source!="EMP"))%>%
  select(SampleID, Station, Latitude, Longitude, SalSurf, Date, Year, IBMR, BPUE)%>%
  group_by(across(-BPUE))%>%
  summarise(BPUE=sum(BPUE), .groups="drop")%>% # Sum each IBMR categories
  st_as_sf(coords=c("Longitude", "Latitude"), crs=4326)%>%
  st_transform(crs=st_crs(deltamapr::R_EDSM_Subregions_Mahardja))%>% # This and next 4 lines to filter data to Suisun Marsh
  st_join(deltamapr::R_EDSM_Subregions_Mahardja%>%
            select(SubRegion))%>%
  st_drop_geometry()%>%
  filter(SubRegion=="Suisun Marsh")%>%
  mutate(doy=yday(Date), #Day of year
         Year_fac=factor(Year), # Factor year for model random effect
         Station_fac=factor(Station), # Factor station for model random effect
         across(c(SalSurf, doy), list(s=~(.x-mean(.x))/sd(.x))), # Center and standardize predictors
         BPUE_log1p=log(BPUE+1)) # log1p transform BPUE for model

```

Set up prediction data for model
```{r}
newdata<-expand_grid(date=mdy(paste(1:12, 15, 2001, sep="/")), # The 15th of each month on a non-leap year
                     SalSurf=seq(min(round(zoop_data_mass$SalSurf), 1), 
                                 round(max(zoop_data_mass$SalSurf), 1), by=0.1))%>% # Salinity sequence nicely rounded to 1 decimal
  mutate(Month=month(date),
         doy=yday(date), # Day of year
         SalSurf_s=(SalSurf-mean(zoop_data_mass$SalSurf))/sd(zoop_data_mass$SalSurf), # center and standardize salinity to match data
         doy_s=(doy-mean(zoop_data_mass$doy))/sd(zoop_data_mass$doy), # center and standardize doy to match data
         Year_fac="2010", # Set arbitrary year (won't be used)
         Station_fac="NZ032")%>% # Set arbitrary station (won't be used)
  select(Month, doy, doy_s, Year_fac, Station_fac, SalSurf, SalSurf_s)
```

model
```{r}
sal_model<-function(group){
  par(mfrow=c(2,2))
  cat("<<<<<<<<<<<<<<<<<<<<<<< modeling", group, ">>>>>>>>>>>>>>>>>>>>>>>>>\n\n")
  
  model<-gam(BPUE_log1p ~ te(SalSurf_s, doy_s, k=c(5,5), bs=c("cs", "cc")) + 
               s(Year_fac, bs="re") + s(Station_fac, bs="re"),
             data=filter(zoop_data_mass, IBMR==group & Year>=2000), 
             method="REML")
  
  cat("-------------gam check-------------\n")
  gam.check(model)
  
  cat("\n\n-------------summary-------------\n")
  print(summary(model))
  
  sal<-predict(model, type="terms", terms="te(SalSurf_s,doy_s)", newdata=newdata, se.fit=T)%>%
    as_tibble()%>%
    mutate(across(everything(), as.vector))%>% # Make everything tidy
    rename(fit=starts_with("fit"), se=starts_with("se.fit"))%>%
    bind_cols(newdata%>% # Add covariate columns before these columns
                select(-Year_fac, -Station_fac, -doy_s, -SalSurf_s), 
              .)
  
  out<-list(model=model, sal=sal)
  return(out)
}
```

Apply model to all groups
```{r}
sal_models<-map(set_names(unique(zoop_data_mass$IBMR)), sal_model)
```

Extract salinity conversions
```{r}
sal_conversions<-map_dfr(sal_models, ~.x[["sal"]], .id = "IBMR")

str(sal_conversions)
```

plot salinity conversions
```{r}
ggplot(sal_conversions, aes(x=SalSurf, y=fit, ymin=fit-se, ymax=fit+se, fill=IBMR))+
  geom_ribbon(alpha=0.4)+
  ylab("fit (log scale)")+
  facet_wrap(~Month)+
  scale_fill_viridis_d()+
  theme_bw()
```

Load in SMSCG modeled salinity
```{r}
SMSCG_sal<-read_csv("Data/SMSCG salinity modeling/bdl_summer_fall_2022_psu_filtered.csv", skip = 10)%>%
  mutate(Month=month(datetime))%>%
  drop_na()%>%
  group_by(Month)%>%
  summarise(across(contains("_bn_"), ~round(mean(.x), 1)))
```

Plot SMSCG modeled salinity
```{r}
ggplot(pivot_longer(SMSCG_sal, cols=contains("_bn_"), names_to="Scenario", values_to="Salinity")%>% # Prepare data for easier plotting
         mutate(Scenario=factor(Scenario, levels=c("bdl_bn_noact", "bdl_bn_smscg6", "bdl_bn_smscg4"))), 
       aes(x=Month, y=Salinity, color=Scenario))+
  geom_line()+
  scale_color_viridis_d(direction = -1)+
  scale_x_continuous(breaks=1:12)+
  theme_bw()+
  theme(legend.position = "bottom")
```


salinity conversion function
```{r}
sal_fix<-function(data, sal_conversions, sal_target, target_column, months){
  
  salinity_conversions<-sal_conversions%>%
    filter(Month%in%months)%>% ## Only choose the months to be included
    mutate(SalSurf=as.character(SalSurf))%>% # Change to character for safer joining
    select(IBMR, Month, SalSurf, fit)
  
  # Are any target salinities outside the range of the model?
  target_outside<-setdiff(as.character(sal_target[[target_column]]), salinity_conversions$SalSurf)
  
  if(length(target_outside)>0){
    cat("Some target salinities are out of range for the prediction. They are:", sort(target_outside))
  }
  
  # select target salinities from the conversions dataset
  sal_conversions_target<-salinity_conversions%>%
    semi_join(sal_target%>%
                mutate(SalSurf=as.character(.data[[target_column]])), 
              by=c("Month", "SalSurf"))%>%
    select(IBMR, Month, fit_target=fit)
  
  if(any(duplicated(paste(sal_conversions_target$Month, sal_conversions_target$IBMR)))){
    stop("sal_conversions were duplicated in filtering out the target salinities")
  }
  
  #Make sure all months are present for each IBMR
  monthcheck<-sal_conversions_target%>%
    group_by(IBMR)%>%
    summarise(N=n_distinct(Month))
  
  if(any(monthcheck$N!=length(months))){
    stop("Some months are missing in the sal_target")
  }
  
  # Set up salinity conversion factors for data and target
  salinity_conversions_correction<-salinity_conversions%>%
    left_join(sal_conversions_target, by=c("IBMR", "Month"))%>% # Join targets to full set of conversions
    mutate(Correction=(exp(fit_target)-1)/(exp(fit)-1))%>% # Calculate correction factor
    select(IBMR, Month, SalSurf, Correction)
  
  # Pull unique salinities for outrange (to test for any data salinities not present in model)
  sals<-unique(salinity_conversions_correction$SalSurf)
  
  # Convert salinity to a character rounded to 1 decimal for safer joining
  data<-data%>%
    mutate(Sal_merge=as.character(round(SalSurf, 1)))
  
  # Are any salinities from the data not present in the model?
  outrange<-setdiff(data$Sal_merge, sals)
  
  if(length(outrange)>0){
    cat("Some salinities are out of range for the prediction. They are:", sort(outrange))
  }
  
  # Join salinity conversions and calculate adjusted BPUE
  data<-data%>%
    left_join(salinity_conversions_correction, by=c("Month", "Sal_merge"="SalSurf", "IBMR"))%>%
    mutate(BPUE_adj=exp(BPUE_log1p+Correction)-1) # Calculate adjusted BPUE and undo the log1p transformation
  
  return(data)
}
```

Prepare data for conversion
```{r}
# From Rosie:
#Calculate water year function
wtr_yr <- function(dates, start_month=12) {
  # Convert dates into POSIXlt
  dates.posix = as.POSIXlt(dates)
  # Year offset
  offset = ifelse(dates.posix$mon == start_month - 1, 1, 0)
  # Water year
  adj.year = dates.posix$year + 1900 + offset
  # Return the water year
  adj.year
}

# Read in water year types
year_types<-read_csv(url("https://raw.githubusercontent.com/rosehartman/DCGmodeling/main/yeartypes.csv"))%>%
  rename(Yearadj = Year)

# Choose the months to be used
months<-c(6,7,8,9,10)

# Filter data to what we want to apply salinity conversions to
zoop_data_mass_bn<-zoop_data_mass%>%
  mutate(Yearadj = wtr_yr(Date),
         Month=month(Date)) %>%
  filter(Year > 2000)%>%
  left_join(year_types) %>%
  filter(Yr_type == "Below Normal" & Month %in% months)%>%
  select(Date, Month, Year, SalSurf, IBMR, BPUE, BPUE_log1p)
```


Apply the conversion function
```{r}
zoop_corrected<-map_dfr(set_names(c("bdl_bn_noact", "bdl_bn_smscg4", "bdl_bn_smscg6")), ~sal_fix(zoop_data_mass_bn, sal_conversions, SMSCG_sal, .x, months), .id="Scenario")%>%
  mutate(Scenario=factor(Scenario, levels=c("bdl_bn_noact", "bdl_bn_smscg6", "bdl_bn_smscg4")))
```

Plot the result
```{r, fig.width=12, fig.height=16}
ggplot(zoop_corrected, aes(x=Date, y=BPUE_adj, color=Scenario))+
  geom_point(alpha=0.4)+
  scale_y_continuous(trans="log1p", breaks = c(0, 10^(1:6)))+
  scale_color_viridis_d(direction = -1, end = 0.9)+
  facet_wrap(~paste(IBMR, Year), scales="free", ncol=5)+
  theme_bw()+
  theme(legend.position = "bottom")
```

Plot BPUE vs BPUE_adj
```{r}
ggplot(zoop_corrected, aes(x=BPUE, y=BPUE_adj, color=Scenario))+
  geom_point(alpha=0.4)+
  scale_y_continuous(trans="log1p", breaks = c(0, 10^(1:6)))+
  scale_x_continuous(trans="log1p", breaks = c(0, 10^(1:6)))+
  scale_color_viridis_d(direction = -1, end = 0.9)+
  facet_wrap(~IBMR, scales="free")+
  theme_bw()+
  theme(legend.position = "bottom")
```

Summarise zoops by IBMR, scenario, and month
```{r}
zoop_bn_sum<-zoop_corrected%>%
  group_by(IBMR, Month, Scenario)%>%
  summarise(BPUE_sum=sum(BPUE_adj), BPUE_sd=sd(BPUE_adj), .groups="drop")

write_csv(zoop_bn_sum, "Outputs/zoop_salinity_corrected_sum.csv")
```

Plot summarized CPUE +/- SD
```{r, fig.width=12, fig.height=12}
ggplot(zoop_bn_sum, aes(x=Month, y=BPUE_sum, ymin=BPUE_sum-BPUE_sd, ymax=BPUE_sum+BPUE_sd, fill=Scenario))+
  geom_pointrange(position = position_dodge(width=0.6), shape=21, color="black", size=1)+
  scale_fill_viridis_d(direction = -1, end = 0.9)+
  facet_wrap(~IBMR, scales="free")+
  theme_bw()+
  theme(legend.position = "bottom")
```

